Biodiversity temporal trends are reshaping food web structure and redundancy in riverine ecosystems

### 2025-04-10 paper/manuscript.Rmd
citation fix

Human pressures are driving profound biodiversity changes, including freshwater
and terrestrial biodiversity declines linked to land degradation
[@newbold_global_2015; @cardinale_is_2018; @danet_past_2024] and reductions in
marine fish biomass due to intensified fishing pressures on predatory species
[@christensen_century_2014]. A central goal of Biodiversity-Ecosystem
Functioning and food web research has been to understand the ecological
consequences of these widespread biodiversity losses for ecosystem functioning
and stability [@loreau_biodiversity_2001; @loreau_ecological_2022;
@thebault_food-web_2003; @schneider_animal_2016; @dunne_network_2002;
@sole_complexity_2001; @allesina_functional_2009; @montoya_press_2009;
@sanders_trophic_2018]. However, research on biodiversity changes remains
largely disconnected from studies examining their ecological impacts, primarily
due to a lack of simultaneous temporal data on biodiversity and food web
dynamics [@danet_species_2021]. As a result, much of our understanding of the
ecological consequences of biodiversity changes over time relies on experimental
and theoretical approaches [@loreau_biodiversity_2001; @loreau_ecological_2022;
@thebault_food-web_2003; @dunne_network_2002; @sole_complexity_2001;
@allesina_functional_2009]. These approaches often rely on the critical
assumption that spatial comparisons of ecological communities can serve as a
proxy for temporal trends—a space-for-time substitution—when direct temporal
data are scarce [@damgaard_a_2019]. To advance our understanding of the
ecological consequences of biodiversity changes, it is essential to investigate
how these changes are reshaping food web structure in natural settings.

>>>>

Human pressures are driving profound biodiversity changes, including freshwater
and terrestrial biodiversity declines linked to land degradation
[@newbold_global_2015; @cardinale_is_2018; @danet_past_2024] and reductions in
marine fish biomass due to intensified fishing pressures on predatory species
[@christensen_century_2014]. A central goal of Biodiversity-Ecosystem
Functioning and food web research has been to understand the ecological
consequences of these widespread biodiversity losses for ecosystem functioning
and stability [@loreau_biodiversity_2001; @loreau_ecological_2022;
@thebault_food-web_2003; @schneider_animal_2016; @dunne_network_2002;
@sole_complexity_2001; @allesina_functional_2009; @montoya_press_2009;
@sanders_trophic_2018]. However, research on biodiversity changes remains
largely disconnected from studies examining their ecological impacts, primarily
due to a lack of simultaneous temporal data on biodiversity and food web
dynamics [@danet_species_2021]. As a result, much of our understanding of the
ecological consequences of biodiversity changes over time relies on experimental
and theoretical approaches [@loreau_biodiversity_2001; @loreau_ecological_2022;
@thebault_food-web_2003; @dunne_network_2002; @sole_complexity_2001;
@allesina_functional_2009]. These approaches often rely on the critical
assumption that spatial comparisons of ecological communities can serve as a
proxy for temporal trends—a space-for-time substitution—when direct temporal
data are scarce [@damgaard_critique_2019]. To advance our understanding of the
ecological consequences of biodiversity changes, it is essential to investigate
how these changes are reshaping food web structure in natural settings.
